Integer Programming Formulations for the Shared Multicast Tree Problem

### 2018-09-05 smt/response.tex

\documentclass[12pt]{article}
\usepackage{a4wide,url}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{color}

\begin{document}
\section*{Response Letter}
We are grateful to the reviewers for their time, effort, and constructive remarks.
All remarks have been accounted for in the revised version, as detailed in our response below.

\section*{Reviewer 2}
Our response in \textcolor{blue}{blue} follows the corresponding remark from the reviewer in \textcolor{red}{red}.

\begin{itemize}
\item \textcolor{red}{The abstract should be shorter.} \textcolor{blue}{
We have revised the abstract, and reduced its length from 22 to 15 lines.
}
\item \textcolor{red}{The literature review has not been organized in a proper format. The references are so old. New related works need to be explained. The authors should use a separate Section for this purpose and compare the novelty of the present approach with them.} \textcolor{blue}{
Following the recommendation of the reviewer, we have reorganized the literature review, and devoted a new section (Section 2 in the revised manuscript)
to this topic. We have also included a review of recent and relevant literature (references ??? in the revision).
In paragraph ??? of Section 2, we now state explicitly the contributions from the present work, and explain how they relate to the current literature.
}
\item \textcolor{red}{They also should consider a separate section on the network model and assumptions.} \textcolor{blue}{
We have restructured the former Section 2 (now Section 3) into three subsections.
The first is devoted to mathematical notation and underlying assumptions, the second addresses the network model,
while the problem definition is given in the third subsection.
}
\item \textcolor{red}{The diagrams of figures are not distinguishable in black and white print. The legends should be corrected.} \textcolor{blue}{
While paying particular attention to their appearance in black and white print, we have given all figures,
including their legends, a complete overhaul.
}
\item \textcolor{red}{All equations should be terminated with "." or "," as well as sentences.} \textcolor{blue}{
Done.
}
\item \textcolor{red}{The legends, markers and other details of the figures are so small.} \textcolor{blue}{
Legends, markers, and other details are now given in larger size.
}
\item \textcolor{red}{A fundamental weak point is that there are no numerical results comparing the proposed scheme with the other new works;} \textcolor{blue}{
We agree that a comparison with alternative contemporary methods largely benefits the paper.
To that end, we have included a comparison with ??? \cite{???}.
Results from the experiments are documented in the new Section 7.?
}
\item \textcolor{red}{The simulation graphs need more expiation and justification.} \textcolor{blue}{
For better justification of the numerical results illustrated graphically in the former Section 6 (now Section 7),
we have added more explanations.
}
\item \textcolor{red}{ The main aim of this paper is the energy, How is the energy issue modeled? Why is not the real models used in the paper?} \textcolor{blue}{
In the new Section 3.2 on the network model, we now explain more detailed how the energy is modeled.
Justification is added in terms of references to relevant and frequently cited works (see [?], ....).
}
\item \textcolor{red}{Overall the novelty of the proposed approach in unclear and the paper is hard to follow
The authors have not considers the wireless problems in the paper.} \textcolor{blue}{
The novelty of and the contributions from the paper are now explicitly stated in Section 2.
We also explain how the paper relates to other works on energy minimization in wireless networks.
}
\end{itemize}

\section*{Reviewer 3}
Our response in \textcolor{blue}{blue} follows the corresponding remark from the reviewer in \textcolor{red}{red}.

\begin{itemize}
\item \textcolor{red}{The paper provides two new ILP formulations for the Shared Multicast Tree (SMT) problem in wireless ad hoc networks.
To address the problem of ILP scalability, the paper also provides a study and proofs of related theorems that shed light on the quality of
solutions achievable by  LP  bounds.
As the number and also provides a new constraint generation scheme.
Numerical results demonstrate the usefulness of the approach. The paper is well written.} \textcolor{blue}{
We are grateful to the reviewer for her/his encouraging remarks, and we appreciate the (s)he finds the paper well written.
}
\item \textcolor{red}{Overall, the methodology makes contribution in the field of optimization as the approach can be used to address other similar problems
that can be modeled as ILP which is not scalable.
However, this paper aims to address the problem of SMT in wireless ad hoc networks and it is
really motivated by the important problem of saving energy in sensor networks.
There I see a certain minor weakness.
The model considered ignores details of energy consumption is sensor network
(see for example Halgamuge et al., Progress in Electromagnetics Research, 12, 2009, or other sources).
It will be good to my opinion to point out this weakness and provide certain qualitative comments to benefit the practitioners,
on how the work can be used in practice, or what is still needed to be done so the work can be useful.} \textcolor{blue}{
We thank the reviewer for suggesting this very useful reference, a citation of which has been included in the revised version of the paper.
We fully agree that there are gaps between the coarse energy model applied in our work and more detailed energy models for energy consumption is sensor networks,
e.g.\ those analyzed by Halgamuge et al.\ (2009).
These gaps, and their possible implications on practical usability of our approaches, are now discussed in the new Section 3.2.
}
\end{itemize}

\noindent
Final comment: We thank the reviewers again for their constructive remarks, which largely helped us to improve the presentation of our findings.

\end{document}

### 2018-09-05 smt/ivanova.tex
MI: Added a reference Halgamuge et. al. (2009), and a paragraph (2nd last paragraph in Sect. 2) that attempts to explain the difference between our energy moddel and the one in the cited paper, and why we do what we do. Label sizes of graphs increased even further
MI: (SMT) Suggestions from Reviewer 2 - Abstract shortened from 22 to 15 lines. Introduced Section 2 Literature overview, although nothing has been changed yet in this section. Section 3 divided into 3 subsections - Assumptions and notation, Network model, Problem definition. The last paragraph in Section 1 (Introduction) is updated

\begin{abstract}
We study the Shared Multicast Tree (SMT) problem in wireless networks.
To support a multicast session between a set of network nodes, SMT aims to establish a wireless 
connection between them, such that the total energy consumption is minimized.
All destinations of the multicast message must be connected, while non-destinations are optional nodes that can be used to relay messages. 
In the well-studied minimum energy multicast (MEM)
problem, the optimal tree depends on the source of the message.
In contrast, SMT asks for a source-independent tree, which substantially reduces the processing overhead at each node.
This means that SMT can be regarded as a Steiner tree problem.
The objective function reflecting power consumption, however, distinguishes SMT clearly from the traditional minimum Steiner tree problem.

In the present work, we develop two 
integer programming formulations for SMT.
Both models are subsequently extended and strengthened by redundant valid inequalities.
Further strengthening is achieved by extending the set of variables and corresponding constraints.
Theorems on relations between the LP bounds corresponding to the models are stated and proved.
As the number of variables in the strongest formulations is a polynomial of degree four in the number of network nodes,
the models are impractical for computing lower bounds in instances beyond a fairly small size.
To make them applicable to larger instances, a constraint generation scheme is developed.
Results from computational experiments with the models demonstrate good promise of the approaches taken.
\keywords{Wireless communication, multicast tree, Steiner tree, LP bound, valid inequalities}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\input{sec1-intro.tex}
\input{sec2-sbtsmt.tex}
\input{sec3-milp.tex}
\input{sec4-relations.tex}
\input{sec5-strong.tex}
\input{sec6-exp.tex}
\input{sec7-conclusion.tex}

@@

>>>>

\begin{abstract}
We study the Shared Multicast Tree (SMT) problem in wireless networks.
To support a multicast session between a set of network nodes, SMT aims to establish a wireless 
connection between them, such that the total energy consumption is minimized.
All destinations of the multicast message must be connected, while non-destinations are optional nodes that can be used to relay messages. 
% MI: I believe these sentences can be disregarded, because the related problems are discussed in detail in the Introduction.
%In the well-studied minimum energy multicast (MEM)
%problem, the optimal tree depends on the source of the message.
%In contrast, SMT asks for a source-independent tree, which substantially reduces the processing overhead at each node.
%This means that SMT can be regarded as a Steiner tree problem.
%The objective function reflecting power consumption, however, distinguishes SMT clearly from the traditional minimum Steiner tree problem.
The objective function reflecting power consumption distinguishes SMT clearly from the traditional minimum Steiner tree problem.

%In the present work,
We develop two integer programming formulations for SMT.
Both models are subsequently extended and strengthened. % by redundant valid inequalities.
%Further strengthening is achieved by extending the set of variables and corresponding constraints.
Theorems on relations between the LP bounds corresponding to the models are stated and proved.
As the number of variables in the strongest formulations is a polynomial of degree four in the number of network nodes,
the models are impractical for computing lower bounds in instances beyond a fairly small size, 
%To make them applicable to larger instances, a constraint generation scheme is developed.
and therefore a constraint generation scheme is developed.
Results from computational experiments with the models demonstrate good promise of the approaches taken.
\keywords{Wireless communication, multicast tree, Steiner tree, LP bound, valid inequalities}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\input{sec1-intro.tex}
\input{sec2-literature.tex}
\input{sec2-sbtsmt.tex}
\input{sec3-milp.tex}
\input{sec4-relations.tex}
\input{sec5-strong.tex}
\input{sec6-exp.tex}
\input{sec7-conclusion.tex}

@@

\bibitem[Halgamuge et al.(2009)]{halgamuge}
Halgamuge MN, Zukerman M, Ramamohanarao K (2009)
An estimation of sensor energy consumption.
Progress In Electromagnetics Research B 12:259--295

### 2018-09-05 smt/sec1-intro.tex
MI: (SMT) Suggestions from Reviewer 2 - Abstract shortened from 22 to 15 lines. Introduced Section 2 Literature overview, although nothing has been changed yet in this section. Section 3 divided into 3 subsections - Assumptions and notation, Network model, Problem definition. The last paragraph in Section 1 (Introduction) is updated according to the new structure of Section 2 and 3.

The combinatorial optimization literature is already rich on Integer Programming (IP) approaches to power minimization in wireless networks.
Since the work of \citet{Wieseltier00onthe}, in particular the MEB and MEM problems have been studied extensively.
Noteworthy contributions to this research include the IP models studied by \citet{das03},
\citet{altinkemer05}, \citet{leggieri08}, \citet{yuan05}, \citet{yuan08}, \citet{bauer08}, and \citet{montemanni11}.
\citet{cagalj02} proved that MEB, and thereby MEM, is NP-hard.
Numerous contributions to fast, but possible inexact methods for MEM have thus appeared, and the interested reader is referred to \citep{hsiao13} for an overview.

The SBT problem has also received some attention in the scientific literature.
It was introduced by \citet{Papadimitriou06SBT}, who also proved that the problem is NP-hard.
In the same article, an approximation algorithm for SBT was developed and analyzed experimentally.
Building on this work, \citet{Haugland12Dual} presented an integer programming model for SBT, along with an associated decomposition approach.

The SBT and SMT problems resemble the more frequently studied \textsc{Range Assignment Problem} (RAP) in the sense that they all ask for an undirected source-independent tree of minimum power.
However, RAP concerns cases where a connected topology must be established by means of \emph{bidirectional} links,
with the feature that power must be assigned to both the sending and the receiving device in each link.
This scenario induces a spanning tree problem, where costs are incurred only for edges that are the most expensive ones incident to any of their end nodes.
Simple instances \citep{Haugland12Dual} show that because of the difference in objective functions, the optimal solutions to RAP and SBT can differ substantially.
\citet{kirousis97} and \citet{clementi99} conduct
hardness results for instances of RAP embedded in three- and two-dimensional Euclidean space, respectively.
Later, IP models have been developed by e.g.\ \citet{althaus03}, \citet{montemanni04}, \citet{das05}, and \citet{Haugland11Compact}.

For the same reasons that MEB has been generalized to MEM, it is natural and desirable to extend SBT to its multicast equivalent.
The \textsc{Minimum Shared Multicast Tree} (SMT) problem is thus at the forefront of this paper,
and neither MEB, MEM, nor RAP will be considered further.
Focus is henceforth on instances in which some of the nodes neither initiate new messages nor need messages from others.
Utilizing such \emph{non-destination} nodes to forward messages can in many cases contribute to substantial reductions in the total power consumption.
Non-destination nodes thus have the potential to play the role of Steiner nodes in the multicast tree.
All devices that can initiate a transmission, referred to as \emph{destination} nodes, also have to receive every message.

Some preliminary work on SMT has recently been published.
\citet{ivanova16isco} introduces the problem, and develops an IP model.
Moreover, the author studies inexact construction and improvement methods, for which computational results are reported.
Contributions from the present work include two sets of IP models for SMT, presented in order of increasing lower bounding capabilities.
By rigorous analysis of the models, we prove relations between the bounds offered by the corresponding continuous relaxations.
Intended for instances beyond the size that can be solved to optimality, we also develop a constraint generation procedure by which strong bounds can be computed.

The remainder of this paper is organized as follows: Section \ref{sec:SBT} describes the SMT problem and gives detailed explanation of its objective function.
Integer linear programming formulations, valid inequalities and their analysis are presented in Section \ref{sec:ILP},
followed by Section \ref{sec:comp}, where the models under study are compared with respect to strength.
In Section \ref{sec:strong}, the models are further extended by variables in higher dimensional spaces,
and a constraint generation procedure is developed.
Computational results are reported in Section \ref{sec:exp},
and suggestions to future work are summarized in Section \ref{sec:conclusion}.

>>>>

Contributions from the present work include two sets of IP models for SMT, presented in order of increasing lower bounding capabilities.
By rigorous analysis of the models, we prove relations between the bounds offered by the corresponding continuous relaxations.
Intended for instances beyond the size that can be solved to optimality, we also develop a constraint generation procedure by which strong bounds can be computed.

The remainder of this paper is organized as follows: 
Section \ref{sec:literature} summarizes relevant literature and previous works.
%Section \ref{sec:SBT} describes the SMT problem and gives detailed explanation of its objective function.
Section \ref{sec:SBT} divided into three subsections introduces the notation, explains the network model, and gives formal definition of the problem.
Integer linear programming formulations, valid inequalities and their analysis are presented in Section \ref{sec:ILP},
followed by Section \ref{sec:comp}, where the models under study are compared with respect to strength.
In Section \ref{sec:strong}, the models are further extended by variables in higher dimensional spaces,
and a constraint generation procedure is developed.
Computational results are reported in Section \ref{sec:exp},
and suggestions to future work are summarized in Section \ref{sec:conclusion}.
